---
title: "Female STEM tertiary graduates"
author: "Sara Berg-Love"
date: "2020-03-14"
output: 
  github_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, warning=FALSE}
# Libraries
library(tidyverse)

# Parameters
file_data <- here::here("data/women_stem_grad.rds")

#===============================================================================

women_stem_grad <- read_rds(file_data)
```

The UNESCO Institute for Statistics (IUS) has a global database to monitor progress on some of their Sustainable Development Goals, such as number 4: equality in education, number 5: gender equality, and number 9: increasing innovation [1][2].

## Summary

```{r summary}
women_stem_grad %>% 
  summary()
```

## 1D data analysis

```{r stem_top_10}
women_stem_grad %>% 
  top_n(10, perct_women_stem_grad) %>% 
  arrange(desc(perct_women_stem_grad))
```

```{r stem_by_year}
women_stem_grad %>% 
  ggplot(aes(year)) +
  geom_bar() +
  scale_x_continuous(breaks = scales::breaks_width(2)) +
  labs(
    title = "The most data is for the years 2015 - 2017",
    x = NULL,
    y = "Number of data points",
    caption = "Source: UNESCO Institute for Statistics"
  )

women_stem_grad %>% 
  count(country) %>% 
  ggplot(aes(n)) +
  geom_bar() +
  scale_x_continuous(breaks = scales::breaks_width(1)) +
  labs(
    title = "Most countries only have 1 data point", 
    x = "Number of data points", 
    y = "Number of countries",
    caption = "Source: UNESCO Institute for Statistics"
  )
```

```{r stem_overall_percentage}
women_stem_grad %>% 
  filter(year == 2017) %>% 
  group_by(country) %>% 
  summarize(perct_women_stem_grad_avg = median(perct_women_stem_grad)) %>% 
  ggplot(aes(perct_women_stem_grad_avg)) +
  geom_histogram(binwidth = 2) +
  geom_vline(
    aes(xintercept = median(perct_women_stem_grad_avg)), color = "blue"
  ) +
  scale_x_continuous(
    breaks = scales::breaks_width(5), 
    labels = scales::percent_format(accuracy = 1, scale = 1)
  ) +
  labs(
    title = 
      "Worldwide the average female STEM tertiary graduates was only 12.5% in 2017",
    x = "Percentage of female STEM graduates",
    y = "Number of data points",
    caption = "Source: UNESCO Institute for Statistics"
  )
```

The most data is for the years 2015 - 2017, but most countries only have 1 data point. It is interesting that Liechtenstein, Gambia, and Myanmar have some of the highest percentages of female tertiary graduates in STEM fields, although it only ranges from 0.6% to 60%. The worldwide average female STEM graduate percentage is very low, only 12.5%.

## Percentage worldwide over time

```{r stem_median_worldwide_percentage}
women_stem_grad %>% 
  ggplot(aes(year, perct_women_stem_grad)) +
  geom_boxplot(aes(group = year), varwidth = TRUE) +
  ggrepel::geom_text_repel(data = . %>% filter(perct_women_stem_grad > 40), aes(label = country), size = 3) +
  labs(
    title = "Median female STEM graduates has not increased over time",
    subtitle = "Although there are increasing numbers of high outliers",
    x = NULL,
    y = "Percentage of female tertiary graduates in STEM fields"
  )
```

Worldwide, the median percentage of female STEM graduates has not increased since 1999. 1998 has a slightly higher percentage, but that could be due to significantly fewer data points. However, there are increasing numbers of outliers on the higher percentage end, which could be a good sign that the female percentage of STEM graduates might increase in the future. There was a slight increase from 2017 to 2018 and not enough data yet for 2019 to make any conclusions.

## Female graduates in STEM fields by country

```{r stem_by_country, out.width = "100%", fig.height=30, message=FALSE}
women_stem_grad %>% 
  group_by(country) %>% 
  filter(length(country) >= 2) %>%
  ungroup() %>% 
  ggplot(aes(year, perct_women_stem_grad)) +
  geom_point(size = 1) +
  geom_line() +
  facet_wrap(vars(country), ncol = 4) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1, scale = 1)) +
  labs(
    title = "Percentage of female tertiary graduates in STEM fields",
    x = "Year",
    y = "Percentage of total tertiary graduates in STEM fields",
    caption = "Source: UNESCO Institute for Statistics"
  )
```

```{r stem_all_countries, out.width = "100%"}
women_stem_grad %>%
  group_by(country) %>% 
  filter(length(country) >= 2) %>%
  ungroup() %>% 
  ggplot(aes(year, perct_women_stem_grad, color = country)) +
  geom_point() +
  geom_line() +
  ggrepel::geom_text_repel(
    aes(label = country), 
    data = . %>% group_by(year) %>% filter(perct_women_stem_grad > 40), 
    nudge_y = 2, 
    size = 3
  ) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1, scale = 1)) +
  theme(legend.position = "none") +
  labs(
    title = 
      "Few countries have female STEM tertiary graduate percentages above 40%",
    x = "Year",
    y = "Percentage of female tertiary graduates in STEM fields",
    caption = "Source: UNESCO Institute for Statistics"
  )
```

Top 5 countries in `r women_stem_grad %>% arrange(year) %>% filter(year == first(year)) %>% distinct(year)` for percentage of female tertiary STEM graduates:
```{r top_5_early}
women_stem_grad %>% 
  arrange(year) %>% 
  filter(year == first(year) + 2) %>% 
  top_n(5, perct_women_stem_grad) %>% 
  arrange(desc(perct_women_stem_grad)) %>% 
  knitr::kable()
```

Bottom 5 countries in `r women_stem_grad %>% arrange(year) %>% filter(year == first(year)) %>% distinct(year)` for percentage of female tertiary STEM graduates:
```{r bottom_5_early}
women_stem_grad %>% 
  arrange(year) %>% 
  filter(year == first(year) + 2) %>% 
  top_n(-5, perct_women_stem_grad) %>% 
  arrange(perct_women_stem_grad) %>% 
  knitr::kable()
```

Top 5 countries in `r women_stem_grad %>% arrange(year) %>% filter(year == last(year) - 1) %>% distinct(year)` for percentage of female tertiary STEM graduates:
```{r top_5_late}
women_stem_grad %>% 
  arrange(year) %>% 
  filter(year == last(year) - 1) %>% 
  top_n(5, perct_women_stem_grad) %>% 
  arrange(desc(perct_women_stem_grad)) %>% 
  knitr::kable()
```

Bottom 5 countries in `r women_stem_grad %>% arrange(year) %>% filter(year == last(year) - 1) %>% distinct(year)` for percentage of female tertiary STEM graduates:
```{r bottom_5_late}
women_stem_grad %>% 
  arrange(year) %>% 
  filter(year == last(year) - 1) %>% 
  top_n(-5, perct_women_stem_grad) %>% 
  arrange(perct_women_stem_grad) %>% 
  knitr::kable()
```


```{r stem_one_country}
plot_women_stem_grad_country <- function(country) {
  women_stem_grad %>% 
    filter(country == {{country}}) %>% 
    ggplot(aes(year, perct_women_stem_grad)) +
    geom_point() +
    geom_line() +
    scale_x_continuous(breaks = scales::breaks_width(width = 1)) +
    scale_y_continuous(
      labels = scales::percent_format(accuracy = 0.1, scale = 1)
    ) +
    labs(
      title = 
        str_glue(
          "Percentage of female tertiary graduates in STEM fields\nin ", 
          {{country}}
        ),
      x = "Year",
      y = "Percentage of total tertiary graduates in STEM fields",
      caption = "Source: UNESCO Institute for Statistics"
    )
}

plot_women_stem_grad_country("United States of America")
plot_women_stem_grad_country("Liechtenstein")
plot_women_stem_grad_country("Gambia")
plot_women_stem_grad_country("Oman")
plot_women_stem_grad_country("Samoa")
```

Pulling out specific outliers and plotting them to see if there are interesting trends, we see that the United States decreased its percentage of female STEM graduates but since 2013 it has been increasing again. Liechtenstein had a stark decrease by almost 50% from 2003 to 2004, which could be wrong data or an unfortunate trend. However, since then it has mostly been increasing. Gambia has had about a 50% increase over 7 years which is good and Oman has increased starkly as well. Samoa has decreased more than 40% in 2 years from 1998 to 2000, but it does not have any updated data since then, so hard to say if that's a general trend.

## Female graduate percentages most and least improved

```{r most_least_improved}
women_stem_grad %>% 
  arrange(year) %>%
  group_by(country) %>%
  summarize(
    change =  
      (last(perct_women_stem_grad) - first(perct_women_stem_grad)) / 
      (last(year) - first(year))
  ) %>% 
  top_n(5, wt = change) %>% 
  arrange(desc(change)) %>% 
  knitr::kable()

women_stem_grad %>% 
  arrange(year) %>%
  group_by(country) %>%
  summarize(
    change =  
      (last(perct_women_stem_grad) - first(perct_women_stem_grad)) / 
      (last(year) - first(year))
  ) %>% 
  top_n(-5, wt = change) %>% 
  arrange(change) %>% 
  knitr::kable()
```

Gambia has increased its percentage of female STEM tertiary graduates by about 41% by year, followed by Brunei Daussalam at 26% per year and Oman at 20% per year. Samoa has decreased its percentage by 16% per year and followed by Antigua and Barbuda and Liechtenstein at 3% per year which are not good trends.

## References

[1] [UNESCO Institute for Statistics.](http://data.uis.unesco.org/)

[2] UNESCO Institute for Statistics. [Data for the Sustainable Development Goals.](http://uis.unesco.org/)
